---
title: "ClickSeq_CYP2D6_doubles_analysis"
author: "Gabrielle Ferra"
date: "2025-02-25"
output: html_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load in libraries, directory, and dataset
```{r load in your libraries, path, and data}
# Load in libraries
library(dplyr)
library(tidyr)
library(ggplot2)

# Load in the path
path = "/Users/gabrielleferra/Desktop/CYP-project/20250828_clickseq_dataset_MINQUAL67_final_datasets_and_analysis/"

# Load the final click-seq DMS activity scores
CYP2D6_data <- read.table(file = file.path(path, "Final_Scores/Pacybara_CYP2D6_activity_scores_singles_fs_and_doubles_12_1e-5_8.231228e-05_2expts_softfilter_MINQUAL67.csv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)
```

### Determine the upper and lower bounds for the scores when you're doing your epistasis analysis - the upper bound is where 95% of synonymous variants have a score less than this value and the lower bound is where 5% of nonsense variants have a score less than this value
```{r determine the score cutoff values}
# Get various synonymous and nonsense stats
synonymous_lowest <- quantile(subset(CYP2D6_data, class == "synonymous")$score, 0.05, na.rm = TRUE)
synonymous_highest <- quantile(subset(CYP2D6_data, class == "synonymous")$score, 0.95, na.rm = TRUE)
nonsense_lowest <- quantile(subset(CYP2D6_data, class == "nonsense" & position < 441)$score, 0.05, na.rm = TRUE)
nonsense_highest <- quantile(subset(CYP2D6_data, class == "nonsense" & position < 441)$score, 0.95, na.rm = TRUE)

nonsense_lowest_2 <- quantile(subset(CYP2D6_data, class == "nonsense")$score, 0.05, na.rm = TRUE)
nonsense_highest_2 <- quantile(subset(CYP2D6_data, class == "nonsense")$score, 0.95, na.rm = TRUE)

# Define the upper and lower bounds
upper_bound <- synonymous_highest
lower_bound <- nonsense_lowest

```

### Apply the bounds
```{r apply the bounds}
# Apply the bounds
CYP2D6_data_doubles <- subset(CYP2D6_data, CYP2D6_data$class == "double")
CYP2D6_data_doubles <- subset(CYP2D6_data_doubles, CYP2D6_data_doubles$score < upper_bound & CYP2D6_data_doubles$score > lower_bound)

CYP2D6_data_singles <- subset(CYP2D6_data, class != "double")
CYP2D6_data_singles <- subset(CYP2D6_data_singles, CYP2D6_data_singles$score < upper_bound & CYP2D6_data_singles$score > lower_bound)

# Separate variant name into variant_1 and variant_2 columns, with their respective scores (score_var_1, score_var_2), mutation classes (class_var_1, class_var_2), and activity classes (activity_class_var_1, activity_class_var_2)
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  # Separate the doubles column into variant_1 and variant_2
  mutate(variant_1 = variant, variant_2 = variant) %>%  # Duplicate variant column
  separate(variant_1, into = c("variant_1", "remove"), sep = ",", extra = "drop") %>%  # Keep first part
  separate(variant_2, into = c("remove", "variant_2"), sep = ",", extra = "drop") %>%  # Keep second part
  mutate(
    variant_2 = trimws(variant_2)  # Remove leading and trailing spaces in variant_2
  ) %>%
  select(-remove) %>%  # Remove the temporary column
  # Join score, class, and activity_class for variant_1
  left_join(
    CYP2D6_data_singles %>%
      select(variant, score, class, activity_class) %>%
      dplyr::rename(
        score_var_1 = score,
        class_var_1 = class,
        activity_class_var_1 = activity_class
      ),
    by = c("variant_1" = "variant")
  ) %>%
  # Join score, class, and activity_class for variant_2
  left_join(
    CYP2D6_data_singles %>%
      select(variant, score, class, activity_class) %>%
      dplyr::rename(
        score_var_2 = score,
        class_var_2 = class,
        activity_class_var_2 = activity_class
      ),
    by = c("variant_2" = "variant")
  )

# omit any NAs
CYP2D6_data_doubles <- na.omit(CYP2D6_data_doubles) # end up with 1181

```


## Compute all of the different doubles scores using the single variant scores
```{r Calculate the scores}
# Get the multiplicative score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(multiplicative_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), 
                                  score_var_1 * score_var_2, 
                                  NA))

# Get the minimum score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(min_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), pmin(score_var_1, score_var_2), NA))

# Get the maximum score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(max_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), pmax(score_var_1, score_var_2), NA))

# Calculate the geometric mean
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(geometric_mean = ifelse(!is.na(score_var_1) & !is.na(score_var_2) & 
                                 score_var_1 > 0 & score_var_2 > 0, 
                                 sqrt(score_var_1 * score_var_2), 
                                 NA))

# Calculate the arithmetic mean
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(arith_mean = ifelse(!is.na(score_var_1) & !is.na(score_var_2), 
                             (score_var_1 + score_var_2) / 2, 
                             NA))

# Calculate the log-transformed additive mean
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(log_additivity = ifelse(!is.na(score_var_1) & !is.na(score_var_2) & score_var_1 > 0 & score_var_2 > 0, 
                                 exp((log(score_var_1) + log(score_var_2)) / 2), 
                                 NA))

# Save this file with the doubles scores + the scores for the respective mutants + all the different types of scores
write.table(CYP2D6_data_doubles, file = file.path(path, "Final_Scores/20251106_Pacybara_CYP2D6_activity_scores_doubles_12_8.231228e-05_2expts_softfilter_MINQUAL67_respectivesingles_allscoretypes.csv"),sep = "\t", quote = FALSE, row.names = F, col.names = T)

```

## Multiplicative score analysis
```{r Multiplicative score analysis}
# load the dataset
CYP2D6_data_doubles <- read.table(file = file.path(path, "Final_Scores/20251106_Pacybara_CYP2D6_activity_scores_doubles_12_8.231228e-05_2expts_softfilter_MINQUAL67_respectivesingles_allscoretypes.csv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)

# Make the scatterplot of the CYP2D6 score vs. the multiplicative score
mult_plot <- ggplot(CYP2D6_data_doubles, aes(x = multiplicative_score, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Multiplicative Score",
    x = "Multiplicative Score",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_multiplicative_syn_and_non_bounds.png"), plot = mult_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_multiplicative_syn_and_non_bounds.pdf"), plot = mult_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ multiplicative_score, data = CYP2D6_data_doubles)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5284907
# w syn and non bounds = 0.4650935

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(difference_mult = score - multiplicative_score)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles$difference_mult, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that on average the click-seq score is higher than the multiplicative score
# Without scores > 1: no longer a significant difference from 0 -> as of now it does appear consistent with the multiplicative model based off of this result!
# With the syn and non bounds: The t-test shows that on average the click-seq score is higher than the multiplicative score. mean difference = 0.0687417 and p-value = 6.55e-15

# I want to look at the variants that have click-seq scores close to the multiplicative score
CYP2D6_data_doubles_small_diff_mult <- CYP2D6_data_doubles %>%
  mutate(diff = abs(score - multiplicative_score)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Minimum score analysis
```{r Minimum score analysis}
# Make the scatterplot of the CYP2D6 score vs. the minimum score
min_plot <- ggplot(CYP2D6_data_doubles, aes(x = min_score, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Minimum Activity Score",
    x = "Minimum Activity Score",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_minimum_syn_and_non_bounds.png"), plot = min_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_minimum_syn_and_non_bounds.pdf"), plot = min_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ min_score, data = CYP2D6_data_doubles)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5299028
# w/ syn and non bounds = 0.4790835 (highest)

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(difference_min = score - min_score)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles$difference_min, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that on average the click-seq score is higher than the minimum score
# Without scores > 1: Still shows that the click-seq score is higher than the minimum on average, which makes sense since we are taking the minimum. the p-value is much higher though (0.001052)
# average of the difference = -0.04352459, p-value = 0.001052, 95% CI [-0.06942735 -0.01762183]; click-seq scores tend to be below the line (lower)
# With syn and non bounds: This t-test shows that on average the click-seq score is higher than the minimum score. the mean difference is 0.06023132 and the p-value = 1.835e-12

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_small_diff_min <- CYP2D6_data_doubles %>%
  mutate(diff = abs(score - min_score)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```


## Maximum score analysis
```{r Maximum score analysis}
# Make the scatterplot of the CYP2D6 score vs. the maximum score
max_plot <- ggplot(CYP2D6_data_doubles, aes(x = max_score, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Maximum Activity Score",
    x = "Maximum Activity Score",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_maximum_syn_and_non_bounds.png"), plot = max_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_maximum_syn_and_non_bounds.pdf"), plot = max_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ max_score, data = CYP2D6_data_doubles)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.09738455
# w/ syn and non bounds = 0.1373625

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(difference_max = score - max_score)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles$difference_max, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that on average the click-seq score is lower than the maximum score
# Without scores > 1: Still shows that the click-seq score is lower than the maximum score on average, but this time the mean difference is much lower (-0.44) with a very significant p-value (2.2e-16)
# With syn and non bounds: This t-test shows that on average the click-seq score is lower than the maximum score with a mean difference of -0.301847 and a p-value of < 2.2e-16

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_small_diff_max <- CYP2D6_data_doubles %>%
  mutate(diff = abs(score - max_score)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Geometric mean analysis
```{r Geometric mean analysis}
# Make the scatterplot of the CYP2D6 score vs. the geometric mean
geom_mean_plot <- ggplot(CYP2D6_data_doubles, aes(x = geometric_mean, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Geometric Mean",
    x = "Geometric Mean",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_geom_mean_syn_and_non_bounds.png"), plot = geom_mean_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_geom_mean_syn_and_non_bounds.pdf"), plot = geom_mean_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ geometric_mean, data = CYP2D6_data_doubles)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5015102
# w/ syn and non bounds = 0.4784046 - higher than multiplicative but lower than minimum

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(difference_geom = score - geometric_mean)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles$difference_geom, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that there is no significant difference between the click-seq scores and the geometric mean, so the variants are randomly distributed above and below the line
# Without scores > 1: Click-seq scores tend to be lower than the geometric mean with a significant p-value (2.2e-16)
# With syn and non bounds: This t-test shows that the click-seq scores tend to be lower than the geometric mean with a mean difference of -0.06569434 and p-value = 2.021e-14

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_small_diff_geom <- CYP2D6_data_doubles %>%
  mutate(diff = abs(score - geometric_mean)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Arithmetic mean analysis
```{r Arithmetic mean analysis}
# Make the scatterplot of the CYP2D6 score vs. the arithmetic mean score
arith_mean_plot <- ggplot(CYP2D6_data_doubles, aes(x = arith_mean, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Arithmetic Mean",
    x = "Arithmetic Mean",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_arith_mean_syn_and_non_bounds.png"), plot = arith_mean_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_arith_mean_syn_and_non_bounds.pdf"), plot = arith_mean_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ arith_mean, data = CYP2D6_data_doubles)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.4486654
# w/ syn and non bounds = 0.4401106

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(difference_arith = score - arith_mean)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles$difference_arith, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that the click-seq score tends to be lower than the additive score
# WIthout scores > 1: Shows the same thing
# With syn and non bounds: This t-test shows that on average the click-seq score is lower than the arithmetic mean with a mean difference of -0.1208079 and p-value < < 2.2e-16

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_small_diff_arith <- CYP2D6_data_doubles %>%
  mutate(diff = abs(score - arith_mean)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Log transformed analysis
```{r Log transformed analysis}
# Make the scatterplot of the CYP2D6 score vs. the log transformed additive mean score
log_additivity_plot <- ggplot(CYP2D6_data_doubles, aes(x = log_additivity, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Log-Transformed Additive Mean",
    x = "Log-Transformed Additive Mean",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_log_mean_syn_and_non_bounds.png"), plot = log_additivity_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_log_mean_syn_and_non_bounds.pdf"), plot = log_additivity_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ log_additivity, data = CYP2D6_data_doubles)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5015102 - same as for arithmetic mean, the acitivty scores do not go up exponentially
# w/ syn and non bounds = 0.4784046 - same as for geometric mean, so higher than multiplicative but lower than minimum, huh

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles <- CYP2D6_data_doubles %>%
  mutate(difference_log = score - log_additivity)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles$difference_log, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that the mean difference is not significantly different from 0. (p = 0.5039, mean = -0.00875, CI = [-0.0345, 0.0169]) show no significant deviation from zero
# Without scores > 1: Shows that the click-seq score is significantly lower than the log mean score
# With syn and non bounds: This t-test shows that on average the click-seq score is lower than the log-transformed additive mean with a mean difference of -0.06569434 and p-value = 2.021e-14

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_small_diff_arith <- CYP2D6_data_doubles %>%
  mutate(diff = abs(score - arith_mean)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

### Filter out doubles and singles that have "possibly" in their activity class
```{r filter out "possibly"'s}

# Begin with running the first 4 chunks of this file to get CYP2D6_data_doubles
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles %>%
  filter(
    !grepl("^possibly", activity_class) &
    !grepl("^possibly", activity_class_var_1) &
    !grepl("^possibly", activity_class_var_2)
  )
```

### Now remake all plots with this subset (no possibly's)
```{r Calculate the scores}
# Get the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(multiplicative_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), 
                                  score_var_1 * score_var_2, 
                                  NA))

# Get the minimum score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(min_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), pmin(score_var_1, score_var_2), NA))

# Get the maximum score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(max_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), pmax(score_var_1, score_var_2), NA))

# Calculate the geometric mean
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(geometric_mean = ifelse(!is.na(score_var_1) & !is.na(score_var_2) & 
                                 score_var_1 > 0 & score_var_2 > 0, 
                                 sqrt(score_var_1 * score_var_2), 
                                 NA))

# Calculate the arithmetic mean
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(arith_mean = ifelse(!is.na(score_var_1) & !is.na(score_var_2), 
                             (score_var_1 + score_var_2) / 2, 
                             NA))

# Calculate the log-transformed additive mean
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(log_additivity = ifelse(!is.na(score_var_1) & !is.na(score_var_2) & score_var_1 > 0 & score_var_2 > 0, 
                                 exp((log(score_var_1) + log(score_var_2)) / 2), 
                                 NA))

# Save this file with the doubles scores + the scores for the respective mutants + all the different types of scores
# w/o possibly's
write.table(CYP2D6_data_doubles, file = file.path(path, "Final_Scores/20251106_Pacybara_CYP2D6_activity_scores_doubles_12_8.231228e-05_2expts_softfilter_MINQUAL67_respectivesingles_allscoretypes_nopossiblys.csv"),sep = "\t", quote = FALSE, row.names = F, col.names = T)

```

## Redo all of the analyses with this subset as well
## Multiplicative score analysis
```{r Multiplicative score analysis}
# Make the scatterplot of the CYP2D6 score vs. the multiplicative score
mult_plot <- ggplot(CYP2D6_data_doubles_filtered, aes(x = multiplicative_score, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Multiplicative Score (filtered)",
    x = "Multiplicative Score",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_multiplicative_syn_and_non_bounds_no_possibly.png"), plot = mult_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_multiplicative_syn_and_non_bounds_no_possibly.pdf"), plot = mult_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ multiplicative_score, data = CYP2D6_data_doubles_filtered)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5284907
# w syn and non bounds = 0.4650935
# w/o possibly: 0.6632713

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(difference_mult = score - multiplicative_score)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles_filtered$difference_mult, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that on average the click-seq score is higher than the multiplicative score
# Without scores > 1: no longer a significant difference from 0 -> as of now it does appear consistent with the multiplicative model based off of this result!
# With the syn and non bounds: The t-test shows that on average the click-seq score is higher than the multiplicative score. mean difference = 0.0687417 and p-value = 6.55e-15
# w/o possibly: no significant difference, mean diff = 0.02632629, p-value = 0.1532

# I want to look at the variants that have click-seq scores close to the multiplicative score
CYP2D6_data_doubles_filtered_small_diff_mult <- CYP2D6_data_doubles_filtered %>%
  mutate(diff = abs(score - multiplicative_score)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Minimum score analysis
```{r Minimum score analysis}
# Make the scatterplot of the CYP2D6 score vs. the minimum score
min_plot <- ggplot(CYP2D6_data_doubles_filtered, aes(x = min_score, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Minimum Activity Score (filtered)",
    x = "Minimum Activity Score",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_minimum_syn_and_non_bounds_no_possibly.png"), plot = min_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_minimum_syn_and_non_bounds_no_possibly.pdf"), plot = min_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ min_score, data = CYP2D6_data_doubles_filtered)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5299028
# w/ syn and non bounds = 0.4790835 (highest)
# w/o possibly = 0.6757666

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(difference_min = score - min_score)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles_filtered$difference_min, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that on average the click-seq score is higher than the minimum score
# Without scores > 1: Still shows that the click-seq score is higher than the minimum on average, which makes sense since we are taking the minimum. the p-value is much higher though (0.001052)
# average of the difference = -0.04352459, p-value = 0.001052, 95% CI [-0.06942735 -0.01762183]; click-seq scores tend to be below the line (lower)
# With syn and non bounds: This t-test shows that on average the click-seq score is higher than the minimum score. the mean difference is 0.06023132 and the p-value = 1.835e-12
# w/o possibly: click-seq tends to be higher, mean diff = 0.04334233, p-value = 0.01727

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_filtered_small_diff_min <- CYP2D6_data_doubles_filtered %>%
  mutate(diff = abs(score - min_score)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Maximum score analysis
```{r Maximum score analysis}
# Make the scatterplot of the CYP2D6 score vs. the maximum score
max_plot <- ggplot(CYP2D6_data_doubles_filtered, aes(x = max_score, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Maximum Activity Score (filtered)",
    x = "Maximum Activity Score",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_maximum_syn_and_non_bounds_no_possibly.png"), plot = max_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_maximum_syn_and_non_bounds_no_possibly.pdf"), plot = max_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ max_score, data = CYP2D6_data_doubles_filtered)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.09738455
# w/ syn and non bounds = 0.1373625
# w/o possibly = 0.1030179

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(difference_max = score - max_score)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles_filtered$difference_max, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that on average the click-seq score is lower than the maximum score
# Without scores > 1: Still shows that the click-seq score is lower than the maximum score on average, but this time the mean difference is much lower (-0.44) with a very significant p-value (2.2e-16)
# With syn and non bounds: This t-test shows that on average the click-seq score is lower than the maximum score with a mean difference of -0.301847 and a p-value of < 2.2e-16
# w/o possibly: click-seq lower, mean difference = -0.457859, p-value < 2.2e-16

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_filtered_small_diff_max <- CYP2D6_data_doubles_filtered %>%
  mutate(diff = abs(score - max_score)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Geometric mean analysis
```{r Geometric mean analysis}
# Make the scatterplot of the CYP2D6 score vs. the geometric mean
geom_mean_plot <- ggplot(CYP2D6_data_doubles_filtered, aes(x = geometric_mean, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Geometric Mean (filtered)",
    x = "Geometric Mean",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_geom_mean_syn_and_non_bounds_no_possibly.png"), plot = geom_mean_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_geom_mean_syn_and_non_bounds_no_possibly.pdf"), plot = geom_mean_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ geometric_mean, data = CYP2D6_data_doubles_filtered)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5015102
# w/ syn and non bounds = 0.4784046 - higher than multiplicative but lower than minimum
# w/o possibly = 0.6244659

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(difference_geom = score - geometric_mean)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles_filtered$difference_geom, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that there is no significant difference between the click-seq scores and the geometric mean, so the variants are randomly distributed above and below the line
# Without scores > 1: Click-seq scores tend to be lower than the geometric mean with a significant p-value (2.2e-16)
# With syn and non bounds: This t-test shows that the click-seq scores tend to be lower than the geometric mean with a mean difference of -0.06569434 and p-value = 2.021e-14
# w/o possibly: click-seq lower, mean diff = -0.09964788, p-value = 2.58e-06

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_filtered_small_diff_geom <- CYP2D6_data_doubles_filtered %>%
  mutate(diff = abs(score - geometric_mean)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Arithmetic mean analysis
```{r Arithmetic mean analysis}
# Make the scatterplot of the CYP2D6 score vs. the arithmetic mean score
arith_mean_plot <- ggplot(CYP2D6_data_doubles_filtered, aes(x = arith_mean, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Arithmetic Mean",
    x = "Arithmetic Mean",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_arith_mean_syn_and_non_bounds_no_possibly.png"), plot = arith_mean_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_arith_mean_syn_and_non_bounds_no_possibly.pdf"), plot = arith_mean_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ arith_mean, data = CYP2D6_data_doubles_filtered)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.4486654
# w/ syn and non bounds = 0.4401106
# w/o possibly = 0.5627399

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(difference_arith = score - arith_mean)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles_filtered$difference_arith, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that the click-seq score tends to be lower than the additive score
# WIthout scores > 1: Shows the same thing
# With syn and non bounds: This t-test shows that on average the click-seq score is lower than the arithmetic mean with a mean difference of -0.1208079 and p-value < < 2.2e-16
# w/o possibly: click-seq is lower, mean diff = -0.2072583, p-value < 2.2e-16

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_filtered_small_diff_arith <- CYP2D6_data_doubles_filtered %>%
  mutate(diff = abs(score - arith_mean)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## Log transformed analysis
```{r Log transformed analysis}
# Make the scatterplot of the CYP2D6 score vs. the log transformed additive mean score
log_additivity_plot <- ggplot(CYP2D6_data_doubles_filtered, aes(x = log_additivity, y = score)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +  # Solid y=x line
  theme_minimal() +
  labs(
    title = "Click-Seq Activity Score vs Log-Transformed Additive Mean",
    x = "Log-Transformed Additive Mean",
    y = "Click-Seq Activity Score"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black", linewidth = 0.3)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_log_mean_syn_and_non_bounds_no_possibly.png"), plot = log_additivity_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_log_mean_syn_and_non_bounds_no_possibly.pdf"), plot = log_additivity_plot, device = "pdf", height =4, width = 8, bg = "white")

# Get the R-squared
lm_fit <- lm(score ~ log_additivity, data = CYP2D6_data_doubles_filtered)
r_squared <- summary(lm_fit)$r.squared
print(r_squared)
# w/o scores > 1 = 0.5015102 - same as for arithmetic mean, the acitivty scores do not go up exponentially
# w/ syn and non bounds = 0.4784046 - same as for geometric mean, so higher than multiplicative but lower than minimum, huh
# w/o possibly = 0.6244659 - same as geom mean??

# Calculate the difference between the click-seq score and the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(difference_log = score - log_additivity)

# Perform a one-sample t-test: This will test if the mean of the differences is significantly different from zero. If the p-value is low (usually below 0.05), it suggests the points are not evenly distributed around the line.
t_test_result <- t.test(CYP2D6_data_doubles_filtered$difference_log, mu = 0)
print(t_test_result)

# With scores > 1: This t-test shows that the mean difference is not significantly different from 0. (p = 0.5039, mean = -0.00875, CI = [-0.0345, 0.0169]) show no significant deviation from zero
# Without scores > 1: Shows that the click-seq score is significantly lower than the log mean score
# With syn and non bounds: This t-test shows that on average the click-seq score is lower than the log-transformed additive mean with a mean difference of -0.06569434 and p-value = 2.021e-14
# w/o possibly: click-seq lower, mean diff = -0.09964788, p-value = 2.58e-06 (same as geom_mean??)

# I want to look at the variants that have click-seq scores close to the minimum score
CYP2D6_data_doubles_filtered_small_diff_arith <- CYP2D6_data_doubles_filtered %>%
  mutate(diff = abs(score - arith_mean)) %>%  # Create new column for absolute difference
  filter(diff < 0.05)

```

## subset to nonsense and synonymous w/o possibly
```{r subset to nonsense and synonymous + something else}
# Begin with running the first 4 chunks of this file to get CYP2D6_data_doubles

# Filter out "possibly" in the activity_class of the double and the singles
# Do we want to do this here?? Starting with this for now, to make sure we are working with the absolute best pickings
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles %>%
  filter(
    !grepl("^possibly", activity_class) &
    !grepl("^possibly", activity_class_var_1) &
    !grepl("^possibly", activity_class_var_2)
  )

# Remove NAs
CYP2D6_data_doubles_filtered <- na.omit(CYP2D6_data_doubles_filtered)

# Calculate the scores
# Get the multiplicative score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(multiplicative_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), 
                                  score_var_1 * score_var_2, 
                                  NA))

# Get the minimum score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(min_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), pmin(score_var_1, score_var_2), NA))

# Get the maximum score
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(max_score = ifelse(!is.na(score_var_1) & !is.na(score_var_2), pmax(score_var_1, score_var_2), NA))

# Calculate the geometric mean
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(geometric_mean = ifelse(!is.na(score_var_1) & !is.na(score_var_2) & 
                                 score_var_1 > 0 & score_var_2 > 0, 
                                 sqrt(score_var_1 * score_var_2), 
                                 NA))

# Calculate the arithmetic mean
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(arith_mean = ifelse(!is.na(score_var_1) & !is.na(score_var_2), 
                             (score_var_1 + score_var_2) / 2, 
                             NA))

# Calculate the log-transformed additive mean
CYP2D6_data_doubles_filtered <- CYP2D6_data_doubles_filtered %>%
  mutate(log_additivity = ifelse(!is.na(score_var_1) & !is.na(score_var_2) & score_var_1 > 0 & score_var_2 > 0, 
                                 exp((log(score_var_1) + log(score_var_2)) / 2), 
                                 NA))

# Subset variants that are synonymous + synonymous 
# do these exist? they do! there are 2 of these doubles, both of them have P489P as one of the variants, and both of them are increased function!! weird..
CYP2D6_data_doubles_filtered_both_syn <- CYP2D6_data_doubles_filtered[CYP2D6_data_doubles_filtered$class_var_1 == "synonymous" & CYP2D6_data_doubles_filtered$class_var_2 == "synonymous", ]

# Classify this subset by adding the "group" column
CYP2D6_data_doubles_filtered_both_syn$group <- "both_syn"

# Subset variants that are nonsense + nonsense
# There are none if we filter out "possibly"s
CYP2D6_data_doubles_filtered_both_non <- CYP2D6_data_doubles_filtered[CYP2D6_data_doubles_filtered$class_var_1 == "nonsense" & CYP2D6_data_doubles_filtered$class_var_2 == "nonsense", ]

# Classify this subset by adding the "group" column
#CYP2D6_data_doubles_filtered_both_non$group <- "both_non"

# Subset variants that are synonymous + nonsense
# CYP2D6_data_doubles_filtered_syn_and_non <- CYP2D6_data_doubles_filtered[
#   (CYP2D6_data_doubles_filtered$class_var_1 == "synonymous" & 
#    CYP2D6_data_doubles_filtered$class_var_2 == "nonsense") |
#   (CYP2D6_data_doubles_filtered$class_var_1 == "nonsense" & 
#    CYP2D6_data_doubles_filtered$class_var_2 == "synonymous"), 
#   ]

# Classify this subset by adding the "group" column
#CYP2D6_data_doubles_filtered_syn_and_non$group <- "syn_and_non"

# Subset variants that are synonymous + missense, frameshift, or nonsense
CYP2D6_data_doubles_filtered_one_syn <- CYP2D6_data_doubles_filtered[
  (CYP2D6_data_doubles_filtered$class_var_1 == "synonymous" & 
   CYP2D6_data_doubles_filtered$class_var_2 %in% c("missense", "frameshift")) |
  (CYP2D6_data_doubles_filtered$class_var_1 %in% c("missense", "frameshift") & 
   CYP2D6_data_doubles_filtered$class_var_2 == "synonymous"), 
  ]

# Classify this subset by adding the "group" column
CYP2D6_data_doubles_filtered_one_syn$group <- "one_syn"

# Subset variants that are nonsense + missense, frameshift, or synonymous
CYP2D6_data_doubles_filtered_one_non <- CYP2D6_data_doubles_filtered[
  (CYP2D6_data_doubles_filtered$class_var_1 == "nonsense" & 
   CYP2D6_data_doubles_filtered$class_var_2 %in% c("missense", "frameshift")) |
  (CYP2D6_data_doubles_filtered$class_var_1 %in% c("missense", "frameshift") & 
   CYP2D6_data_doubles_filtered$class_var_2 == "nonsense"), 
  ]

# Classify this subset by adding the "group" column
CYP2D6_data_doubles_filtered_one_non$group <- "one_non"

# Combine all of these subsets together 
CYP2D6_data_doubles_filtered_combined <- rbind(
  CYP2D6_data_doubles_filtered_both_syn,
  #CYP2D6_data_doubles_filtered_both_non,
  #CYP2D6_data_doubles_filtered_syn_and_non,
  CYP2D6_data_doubles_filtered_one_syn,
  CYP2D6_data_doubles_filtered_one_non
)

# Make the scatterplot of the CYP2D6 score vs. the multiplicative score
subset_mult_plot <- ggplot(
  CYP2D6_data_doubles_filtered_combined,
  aes(x = multiplicative_score, y = score, color = group)
) +
  geom_point(alpha = 0.8, size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +
  theme_minimal() +
  labs(
    title  = "Click-Seq Score vs Multiplicative Score for Synonymous and Nonsense Variants",
    x      = "Multiplicative Score",
    y      = "Click-Seq Activity Score"
  ) +
  scale_color_manual(
    values = c(
      "both_syn"     = "blue",
      "one_syn"      = "#A353C2",
      "one_non"      = "red"
    ),
    labels = c(
      "both_syn"     = "Both Synonymous",
      "one_syn"      = "One Synonymous",
      "one_non"      = "One Nonsense"
    ),
    breaks = c("both_syn", "one_syn", "one_non"),  # <- legend order
    name = "Variant Types"
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 3.5, alpha = 0.8))
    # size = 5 makes the dots in the legend bigger
    # alpha = 1 makes sure they’re not faint in the legend
  ) +
  theme(
    plot.title   = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title   = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 10),
    axis.line    = element_line(color = "black", linewidth = 0.3),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12)
  )

print(subset_mult_plot)


# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_multiplicative_syn_and_non_bounds_no_possibly_control_set.png"), plot = subset_mult_plot, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_clickseq_v_multiplicative_syn_and_non_bounds_no_possibly_control_set.pdf"), plot = subset_mult_plot, device = "pdf", height =4, width = 8, bg = "white")

```

## Subsets of nonsense and synonymous w/ possibly
```{r subset to nonsense and synonymous + something else}
# Begin with running the first 4 chunks of this file to get CYP2D6_data_doubles

# Remove NAs
CYP2D6_data_doubles <- na.omit(CYP2D6_data_doubles)

# Subset variants that are synonymous + synonymous 
# do these exist? they do! there are 2 of these doubles, both of them have P489P as one of the variants, and both of them are increased function!! weird..
CYP2D6_data_doubles_both_syn <- CYP2D6_data_doubles[CYP2D6_data_doubles$class_var_1 == "synonymous" & CYP2D6_data_doubles$class_var_2 == "synonymous", ]

# Classify this subset by adding the "group" column
CYP2D6_data_doubles_both_syn$group <- "both_syn"

# Subset variants that are nonsense + nonsense
# There are none if we filter out "possibly"s
CYP2D6_data_doubles_both_non <- CYP2D6_data_doubles[CYP2D6_data_doubles$class_var_1 == "nonsense" & CYP2D6_data_doubles$class_var_2 == "nonsense", ]

# Classify this subset by adding the "group" column
#CYP2D6_data_doubles_both_non$group <- "both_non"

# Subset variants that are synonymous + nonsense
CYP2D6_data_doubles_syn_and_non <- CYP2D6_data_doubles[
  (CYP2D6_data_doubles$class_var_1 == "synonymous" &
   CYP2D6_data_doubles$class_var_2 == "nonsense") |
  (CYP2D6_data_doubles$class_var_1 == "nonsense" &
   CYP2D6_data_doubles$class_var_2 == "synonymous"),
  ]

# Classify this subset by adding the "group" column
CYP2D6_data_doubles_syn_and_non$group <- "syn_and_non"

# Subset variants that are synonymous + missense, frameshift, or nonsense
CYP2D6_data_doubles_one_syn <- CYP2D6_data_doubles[
  (CYP2D6_data_doubles$class_var_1 == "synonymous" & 
   CYP2D6_data_doubles$class_var_2 %in% c("missense", "frameshift")) |
  (CYP2D6_data_doubles$class_var_1 %in% c("missense", "frameshift") & 
   CYP2D6_data_doubles$class_var_2 == "synonymous"), 
  ]

# Classify this subset by adding the "group" column
CYP2D6_data_doubles_one_syn$group <- "one_syn"

# Subset variants that are nonsense + missense, frameshift, or synonymous
CYP2D6_data_doubles_one_non <- CYP2D6_data_doubles[
  (CYP2D6_data_doubles$class_var_1 == "nonsense" & 
   CYP2D6_data_doubles$class_var_2 %in% c("missense", "frameshift")) |
  (CYP2D6_data_doubles$class_var_1 %in% c("missense", "frameshift") & 
   CYP2D6_data_doubles$class_var_2 == "nonsense"), 
  ]

# Classify this subset by adding the "group" column
CYP2D6_data_doubles_one_non$group <- "one_non"

# Combine all of these subsets together 
CYP2D6_data_doubles_combined <- rbind(
  CYP2D6_data_doubles_both_syn,
  #CYP2D6_data_doubles_both_non,
  CYP2D6_data_doubles_syn_and_non,
  CYP2D6_data_doubles_one_syn,
  CYP2D6_data_doubles_one_non
)

# Make the scatterplot of the CYP2D6 score vs. the multiplicative score
subset_mult_plot <- ggplot(
  CYP2D6_data_doubles_combined,
  aes(x = multiplicative_score, y = score, color = group)
) +
  geom_point(alpha = 0.8, size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "#9C0520", linewidth = 0.7) +
  theme_minimal() +
  labs(
    title  = "Click-Seq Score vs Multiplicative Score for Synonymous and Nonsense Variants",
    x      = "Multiplicative Score",
    y      = "Click-Seq Activity Score"
  ) +
  scale_color_manual(
    values = c(
      "both_syn"     = "blue",
      "one_syn"      = "#BC13FE",
      "syn_and_non"  =  "#FF5F1F",
      "one_non"      = "red",
      "both_non"     = "#A10303"
    ),
    labels = c(
      "both_syn"     = "Both Synonymous",
      "one_syn"      = "One Synonymous",
      "syn_and_non"  = "Synonymous + Nonsense",
      "one_non"      = "One Nonsense",
      "both_non"     = "Both Nonsense"
    ),
    breaks = c("both_syn", "one_syn", "syn_and_non", "one_non", "both_non"),  # <- legend order
    name = "Variant Types"
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 3.5, alpha = 0.8))
  ) +
  theme(
    plot.title   = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title   = element_text(size = 12, face = "bold"),
    axis.text    = element_text(size = 10),
    axis.line    = element_line(color = "black", linewidth = 0.3),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text  = element_text(size = 12),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA)
  )

print(subset_mult_plot)

##F2672C" old syn_and_non
##A353C2" old one_syn


# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251224_clickseq_v_multiplicative_syn_and_non_bounds_control_set.png"), plot = subset_mult_plot, device = "png", height =4, width = 8)
ggsave(file = file.path(path, "Final_Plots/20251224_clickseq_v_multiplicative_syn_and_non_bounds_control_set.pdf"), plot = subset_mult_plot, device = "pdf", height =4, width = 8)

```

### Plot the distributions of activity class for our control set
```{r activity class distribution}
# Many of the doubles were assigned an increased activity class based on their click-seq result so now I want to look at the activity class distribution of our control set made up of synonymous + something and nonsense + something doubles - but I want to plot these distributions by the groups

# Set the order for activity_class
CYP2D6_data_doubles_filtered_combined$activity_class <- factor(CYP2D6_data_doubles_filtered_combined$activity_class,levels = c("nonsense-like", "decreased", "wt-like", "increased"))


# Bar plot for all groups together
subset_bar_plot_all <- ggplot(CYP2D6_data_doubles_filtered_combined, aes(x = activity_class, fill = activity_class)) +
  geom_bar(aes(y = ..count..), stat = "count") +  # Count the number of variants
  theme_minimal() +
  labs(
    title = "Activity Class Distribution for All Groups",
    x = "Activity Class",
    y = "Count of Variants"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "none"  # Remove the legend
  )

# Save the plot
ggsave(file = file.path(path, "Plots/20250324_clickseq_activity_class_bar_plot_syn_and_non_bounds_no_possibly_control_set.png"), plot = subset_bar_plot_all, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Plots/20250324_clickseq_activity_class_bar_plot_syn_and_non_bounds_no_possibly_control_set.pdf"), plot = subset_bar_plot_all, device = "pdf", height =4, width = 8, bg = "white")

# Bar plot by activity_class and group
subset_bar_plot_group <- ggplot(CYP2D6_data_doubles_filtered_combined, aes(x = activity_class)) +
  geom_bar(aes(y = ..count.., fill = activity_class), stat = "count") +  # Count the number of variants
  facet_wrap(~ group) +  # Create separate plots for each group
  theme_minimal() +
  labs(
    title = "Activity Class Distribution by Group",
    x = "Activity Class",
    y = "Count of Variants"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "none",  # Remove the legend
    strip.background = element_rect(color = "black", fill = "white", size = 1),  # Add box around each facet
    strip.text = element_text(size = 12),  # Ensure facet labels are visible
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
  )

# Save the plot
ggsave(file = file.path(path, "Plots/20250324_clickseq_activity_class_bar_plot_syn_and_non_bounds_no_possibly_control_set_by_group.png"), plot = subset_bar_plot_group, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Plots/20250324_clickseq_activity_class_bar_plot_syn_and_non_bounds_no_possibly_control_set_by_group.pdf"), plot = subset_bar_plot_group, device = "pdf", height =4, width = 8, bg = "white")

```

### Plot the distributions of the estimated scores for the multiplicative score vs. the click-seq score
### Load required packages for plots
```{r Load_packages, message=FALSE, warning = FALSE}
library(ggplot2)
library(plyr)
library(tidyr)
library(reshape)
library(scales)
library(GGally)
library(grid)
library(gridExtra)
library(zoo)
library(cowplot)

## The package versions used by the author as of 01/12/2023
packageVersion("ggplot2")     #‘3.4.0’
packageVersion("plyr")     #‘1.8.8’
packageVersion("tidyr")     #‘1.2.1’
packageVersion("reshape")     #‘0.8.9’
packageVersion("scales")     #‘1.2.1’
packageVersion("GGally")     #‘2.1.2’
packageVersion("grid")     #‘4.2.2’
packageVersion("gridExtra")     #2.3
packageVersion("zoo")     #1.8.11
packageVersion("cowplot")     #1.1.1

```

### Functions to plot score correlation matrix
```{r score_correlation, warning=FALSE} 
# Scatterplot function
scatterplot_correlation_fxn <- function(data, mapping, ...) {

  xvar <- rlang::as_label(mapping$x)
  yvar <- rlang::as_label(mapping$y)

  ggplot(data = data, aes_string(x = yvar, y = xvar)) +
    geom_point(alpha = 0.6, ...) +
    scale_x_continuous(
      breaks = c(0, 0.5, 1.0),
      labels = c("0.0", "0.5", "1.0")
    ) +
    scale_y_continuous(
      breaks = c(0, 0.5, 1.0),
      labels = c("0.0", "0.5", "1.0")
    ) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    )
}

histogram_fxn <- function(data, mapping, ...){
  # Set the desired order of the class factor levels
  data$class <- factor(data$class, levels = c("double", "missense", "frameshift", "nonsense", "synonymous"))
  
  p <- ggplot(data = data, mapping = mapping) + 
    geom_histogram(aes(fill=class), binwidth=0.15, alpha = 0.5, size=0.2, position="stack", color="black") +
    scale_fill_manual(values=c("frameshift" = "#FF7518", "missense"="grey", "nonsense"="red", "synonymous"="blue", "double"="#FFDB58")) + 
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_blank(), 
          axis.line = element_line(colour = "black"))
  p
}
```

### Set up variables for the score correlation matrix 
```{r Choose score correlation matrix variables}

# Choose which replicates you want to compare
a = "score"
b = "multiplicative_score"

```

### Plot and save the score correlation matrix
```{r}
# Create the dataset with the desired scores_______________________________________________
#cyp_variants_min <- CYP2D6_data_doubles_filtered[,c("class", a, b)]
cyp_variants_min <- CYP2D6_data_doubles[,c("class", a, b)]

# Designate the file name
end <- "clickseq_vs_multiplicative_score_distributions"

# Calculate the Pearson correlation between the scores
cyp_pearson_cor_matrix <- as.matrix(cor(cyp_variants_min[,c(a,b)], use="pairwise.complete.obs", method = "pearson")) ###pairwise.complete.obs uses the non-NA values when calculating the correlation between score1 and score2  , Pearson : the linear relationship between two continuous variables

# Calculate the Spearman correlation between the scores
cyp_spearman_cor_matrix <- as.matrix(cor(cyp_variants_min[,c(a,b)], use="pairwise.complete.obs", method = "spearman")) ### Spearman : the monotonic relationship between two continuous or ordinal variables

# Plot the correlation between the two distributions of activity scores
library(GGally)
library(ggplot2)

cyp_scatterplot_matrix_pearson <-
  ggpairs(
    cyp_variants_min,
    columns = c(a, b),
    lower = list(continuous = scatterplot_correlation_fxn),
    diag  = list(continuous = histogram_fxn),
    columnLabels = c("Click-seq Activity Score", "Multiplicative Activity Score")
  ) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold")
  )

# Save the plots as .png and .pdf 
ggsave(file = file.path(path, paste("Final_Plots/20251225_", end, ".png", sep = "")), plot = cyp_scatterplot_matrix_pearson, device = "png", height = 14, width = 14, units = "cm")
ggsave(file = file.path(path, paste("Final_Plots/20251225_", end, ".pdf", sep = "")), plot = cyp_scatterplot_matrix_pearson, device = "pdf", height = 14, width = 14, units = "cm")

```

### Plot the spatial distribution of these double mutants for no possibly's 
```{r notes}
# Extract the number from variant_1 and store it in var_1_pos
CYP2D6_data_doubles_filtered$var_1_pos <- as.numeric(sub("^[A-Za-z](\\d+)[A-Za-z]$", "\\1", CYP2D6_data_doubles_filtered$variant_1))

# Extract the number from variant_2 and store it in var_2_pos
CYP2D6_data_doubles_filtered$var_2_pos <- as.numeric(sub("^[A-Za-z](\\d+)[A-Za-z]$", "\\1", CYP2D6_data_doubles_filtered$variant_2))

# Scatterplot of var_1_pos vs. var_2_pos
spatial_dist <- ggplot(CYP2D6_data_doubles_filtered, aes(x = var_1_pos, y = var_2_pos)) +
  geom_point(alpha = 0.6) +  # Scatter plot with semi-transparent points
  theme_minimal() +  # Minimal theme for the plot
  labs(
    title = "Spatial Distribution of Double Mutants (Filtered: No Possibly's)",
    x = "Position of Variant 1",
    y = "Position of Variant 2"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10)
  )

# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_spatial_distribution_of_double_mutants_no_possibly.png"), plot = spatial_dist, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_spatial_distribution_of_double_mutants_no_possibly.pdf"), plot = spatial_dist, device = "pdf", height =4, width = 8, bg = "white")

```

### Plot the spatial distribution of these double mutants including the possibly's (still bounded by synonymous and nonsense)
```{r notes}
# Extract the number from variant_1 and store it in var_1_pos
CYP2D6_data_doubles$var_1_pos <- as.numeric(sub("^[A-Za-z](\\d+)[A-Za-z]$", "\\1", CYP2D6_data_doubles$variant_1))

# Extract the number from variant_2 and store it in var_2_pos
CYP2D6_data_doubles$var_2_pos <- as.numeric(sub("^[A-Za-z](\\d+)[A-Za-z]$", "\\1", CYP2D6_data_doubles$variant_2))

# Scatterplot of var_1_pos vs. var_2_pos
spatial_dist <- ggplot(CYP2D6_data_doubles, aes(x = var_1_pos, y = var_2_pos)) +
  geom_point(alpha = 0.6) +
  theme_minimal() +
  labs(
    title = "Spatial Distribution of Double Mutants",
    x = "Position of Variant 1",
    y = "Position of Variant 2"
  ) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      size = 14,
      face = "bold",
      margin = margin(b = 12)  # space between title and panel
    ),
    axis.title.x = element_text(
      size = 12,
      face = "bold",
      margin = margin(t = 10)  # space between x label and panel
    ),
    axis.title.y = element_text(
      size = 12,
      face = "bold",
      margin = margin(r = 10)  # space between y label and panel
    ),
    axis.text  = element_text(size = 10),
    # <- this draws the visible axes
    axis.line  = element_line(color = "black", linewidth = 0.3),
    # optional: lighten the grid so axes pop more
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank()
  )

print(spatial_dist)


# Save the plot
ggsave(file = file.path(path, "Final_Plots/20251106_spatial_distribution_of_double_mutants.png"), plot = spatial_dist, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Final_Plots/20251106_spatial_distribution_of_double_mutants.pdf"), plot = spatial_dist, device = "pdf", height =4, width = 8, bg = "white")

```

### Plot the activity class distribution for synonymous and nonsense singles
```{r activity class distribution}
# Load in the path
path = "/Users/gabrielleferra/Desktop/CYP-project/20250115/20250224_doubles_score_analysis"

# Load in your data
CYP2D6_data <- read.table(file = file.path(path, "Scores/Pacybara_CYP2D6_activity_scores_singles_fs_and_doubles_124_1e-5_5681e-5_3expts.csv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)

# Subset to synonymous and nonsense singles
CYP2D6_data_syn <- subset(CYP2D6_data, CYP2D6_data$class == "synonymous")
CYP2D6_data_non <- subset(CYP2D6_data, CYP2D6_data$class == "nonsense")

# Set the order for activity_class
CYP2D6_data_syn$activity_class <- factor(CYP2D6_data_syn$activity_class,levels = c("nonsense-like", "possibly_nonsense-like", "decreased", "possibly_decreased", "wt-like", "possibly_wt-like", "increased", "possibly_increased"))
CYP2D6_data_non$activity_class <- factor(CYP2D6_data_non$activity_class,levels = c("nonsense-like", "possibly_nonsense-like", "decreased", "possibly_decreased", "wt-like", "possibly_wt-like", "increased", "possibly_increased"))

# Bar plot for syn singles together
subset_bar_plot_syn <- ggplot(CYP2D6_data_syn, aes(x = activity_class, fill = activity_class)) +
  geom_bar(aes(y = ..count..), stat = "count") +  # Count the number of variants
  theme_minimal() +
  labs(
    title = "Activity Class Distribution for Synonymous Singles",
    x = "Activity Class",
    y = "Count of Variants"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "none"  # Remove the legend
  )

# Save the plot
ggsave(file = file.path(path, "Plots/20250408_clickseq_activity_class_bar_plot_syn.png"), plot = subset_bar_plot_syn, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Plots/20250408_clickseq_activity_class_bar_plot_syn.pdf"), plot = subset_bar_plot_syn, device = "pdf", height =4, width = 8, bg = "white")

# Bar plot for nonsense singles together
subset_bar_plot_non <- ggplot(CYP2D6_data_non, aes(x = activity_class, fill = activity_class)) +
  geom_bar(aes(y = ..count..), stat = "count") +  # Count the number of variants
  theme_minimal() +
  labs(
    title = "Activity Class Distribution for Nonsense Singles",
    x = "Activity Class",
    y = "Count of Variants"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "none"  # Remove the legend
  )

# Save the plot
ggsave(file = file.path(path, "Plots/20250408_clickseq_activity_class_bar_plot_non.png"), plot = subset_bar_plot_non, device = "png", height =4, width = 8, bg = "white")
ggsave(file = file.path(path, "Plots/20250408_clickseq_activity_class_bar_plot_non.pdf"), plot = subset_bar_plot_non, device = "pdf", height =4, width = 8, bg = "white")

```

## Plot activity score ditributions with multiplicative and minimum score distributions plotted on top
```{r activity class distribution}
# Load combined activity score data
CYP2D6_data <- CYP2D6_data_doubles # length is 1,671 - we have data for both singles and the double for this many doubles
end = "doubles_12_5681e-5"
i = 350
#a = "no alleles" # Uncomment if you don't want to include the alleles of interest 
                 # the single aa changes cannot be included once you have subsetted the data to just doubles
                 # feel free to include double variants of interest

#__________________________________________________________________________________________

# base histogram
library(dplyr)
library(ggplot2)

# Filter to the same data used in the histogram (exclude wt)
dat_hist <- subset(CYP2D6_data, class != "wt")

binwidth_val <- 0.05

# how many variants go into each model curve
n_score <- nrow(dat_hist)
n_mult  <- sum(!is.na(dat_hist$multiplicative_score))
n_min   <- sum(!is.na(dat_hist$min_score))

hist_plot <- ggplot() +
  # histogram
  geom_histogram(
    data = dat_hist,
    aes(x = score, fill = class),
    binwidth = binwidth_val,
    alpha = 0.5,
    size = 0.25,
    position = "stack",
    color = "black"
  ) +

  # multiplicative model line (solid black)
  geom_density(
    data = dat_hist,
    aes(
      x = multiplicative_score,
      y = after_stat(..density.. * n_mult * binwidth_val),
      linetype = "Multiplicative model",
      color    = "Multiplicative model"
    ),
    linewidth = 0.8,
    alpha = 0.4
  ) +

  # # minimum model line (dashed gray)
  # geom_density(
  #   data = dat_hist,
  #   aes(
  #     x = min_score,
  #     y = after_stat(..density.. * n_min * binwidth_val),
  #     linetype = "Minimum model",
  #     color    = "Minimum model"
  #   ),
  #   linewidth = 0.8,
  #   alpha = 0.4
  # ) +

  scale_fill_manual(
    values = c(
      "frameshift" = "#FF7518",
      "missense"   = "grey",
      "nonsense"   = "red",
      "synonymous" = "blue",
      "double"     = "#FFDB58"
    ),
    labels = c(
      "missense"   = "Missense",
      "frameshift" = "Frameshift",
      "nonsense"   = "Nonsense",
      "synonymous" = "Synonymous",
      "double"     = "Double Mutants"
    ),
    guide = "none"
  ) +

  scale_color_manual(
    values = c(
      "Multiplicative model" = "black",
      "Minimum model"        = "grey40"
    ),
    breaks = c("Multiplicative model", "Minimum model"),
    name = "Model"
  ) +

  scale_linetype_manual(
    values = c(
      "Multiplicative model" = "solid",
      "Minimum model"        = "dashed"
    ),
    breaks = c("Multiplicative model", "Minimum model"),
    name = "Model"
  ) +

  theme_classic() +
  ggtitle("Click-seq Activity Score Distribution vs Predictive Models for Double Variants") +
  xlab("Activity Score") +
  ylab("Count") +
  coord_cartesian(ylim = c(0, 80)) +
  theme(
    axis.text.x   = element_text(color = "black", size = 12),
    axis.text.y   = element_text(color = "black", size = 12),
    axis.title.x  = element_text(size = 14, face = "bold"),
    axis.title.y  = element_text(size = 14, face = "bold"),
    plot.title    = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.position = "none"  ,   # ⬅ turn off legend completely
    legend.title  = element_text(size = 14, face = "bold"),
    legend.text   = element_text(size = 12)
  )

print(hist_plot)




# Save the histogram plot
ggsave(file = file.path(path, paste("Final_Plots/20251106_double_mutant_score_distribution_vs_mult.png", sep = "")), plot = hist_plot, device = "png", height =3.5, width = 5)

ggsave(file = file.path(path, paste("Final_Plots/20251106_double_mutant_score_distribution_vs_mult.pdf", sep = "")), plot = hist_plot, device = "pdf", height =3.5, width = 5)

```
